Analysis

### 2021-01-15 analysis.Rmd
proofed.

Workflow and analysis

First import data (see functions)

@@

Inspect the data, and note that many columns data have been transformed from their raw form.

@@

Ask Suzanne for raw data so that we can, e.g. model church frequency as a monotonic categorical variable, and pick appropriate sampling distributions for the outcome variables.

@@

```{r, include = FALSE}
# get columns for different indicators 
rel <- get_rel_vars(df)
wbgen <- get_wb_gen(df)
wbphys <- get_wb_phys(df)
wbpsych <- get_wb_psych(df)
wbsoc <- get_wb_soc(df)
head(wbsoc)
head(rel)
```

Exploration Graphs.  Inspect the data.

@@

@@

# Specify our causal model

@@

Which variables should we condition on to understand the relationship between religion and well-being?

@@

@@

## The hard work

@@

There are many theories of how religion affects well-being we might consider (cite Highland et al). We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.

@@

```{r}
tidy_ggdag2 <- dagify(
  wb  ~ support + ses + male + country + U2,
  support ~ church + male + country,
  church ~ age + male + country + U1,
  ses ~ country + age + U2,
  exposure =  "church",
  outcome =   "wb"
) %>%
  tidy_dagitty()

@@

Before proceeding with the model, let's note a curiosity. If our  interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias!

@@

Let this serve as a warning for us to think about our causation, and what we might want to know before we leap into modeling

## The model

@@

## degroup variables for within and between analysis
## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
d1 <-
  cbind(d1,
        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
head(d1)
```

@@

We can see there is no effect of church attendance on anxiety (according to this model). Note that throughout, the plots will sample 300 draws from the posterior distribution.

@@

Before racing ahead, lets do some posterior predictive checks

@@

When we do a posterior predictive check we find problems in the predicted distribution. This we called this a gaussian, but it is not. The data have large peaks reflecting the categorical qualities of the outcome variable.

We should work with the original, untransformed values of the data.

@@

<!-- ```{r cache = TRUE} -->
<!-- plot(ggeffects::ggpredict(m1_prior, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +  scale_fill_brewer(palette= "Set2") + labs( title = "Predicted effects of church on anxiety") -->
<!-- ``` -->

or we can simulate and see that our prior is scattered across five standard deviations of anxiety, and we can see our prior is telling us that anything can happen

@@

Nothing happening at the individual level.

```{r include = FALSE}
# graph not as nice
# plot(ggeffects::ggpredict(m1, terms = c("ch_s[minmax]"),type = "fe", ci.lvl = 0.89))  +
#    scale_fill_brewer(palette= "Set2") + labs(
#     title = "Predicted effects of church on anxiety")
```

But lots of variability at the country level intercepts for anxiety. We can see that India is over two standard deviations  less anxious than the average culture level anxiety. By contrast, Japan, Israel, and France are over .2 standard deviations more anxious than average. There looks to be perhaps some covariance between the country intercepts and slopes... maybe though all estimates crosses zero. We'd either want to use stronger priors or collect data from more countries than are in this sample.

@@

Let's delving into this question of country level effects, but asking whether people who are relatively more church attending tend to have more anxiety.

@@

- between_church centering = the adjustment for country level differences in church attendance.

@@

@@

No reliable effect within,

@@

But a between-country effect emerges:

@@

Same workflow as before.

>>>>

## Workflow and analysis

First, we import the data (see functions).

@@

Next, we inspect the data. A key problem is that many columns data have been transformed from their raw form into a number between 0 and 1. Later we'll see that these transformations render the modelling less reliable than if the data were retained.

So for the next iteration, we'll need the data as they were recorded. Specially we'll focus on the variables such as

> How often do you have negative feelings such as blue mood, despair, anxiety, depression

We'll investigate church frequency and belief in God as predictors of this dimension of well-being.

@@

@@

```{r, include = FALSE}
# get columns for different indicators 
rel <- get_rel_vars(df)
wbgen <- get_wb_gen(df)
wbphys <- get_wb_phys(df)
wbpsych <- get_wb_psych(df)
wbsoc <- get_wb_soc(df)

Next some exploration Graphs.  Inspecting the sample responses on religion.

@@

India is a big outlier: much more religious than other countries.

Let's look at the anxiety indicator: india is much lower:
. 
```{r}
wb1 <- wbpsych %>%
  ggplot(aes(
    x = country,
    y = wb_psychANXIETY,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  ggtitle("Importance Believe in God") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
wb1

We're interested in  meaning of life as a causal mediator of anxiety

```{r}
wb2 <- wbpsych %>%
  ggplot(aes(
    x = country,
    y = wb_psychMEANING,
    group = country,
    color = country
  )) +
  geom_boxplot(notch = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  geom_point(position = "jitter", alpha = .1) +
  ggtitle("Importance Believe in God") +
  ylab("") + 
  theme(axis.text.x = element_text(angle = 30))
wb2
```

India again very high, and France is very low.  Can we really assume measurement invariance across these countries?  Are people so really different? As with any model, we condition on assumptions, however at another iteration of this study, however for now we will assume measurement invariance.

@@

## A critical step: the causal model

@@

### Which variables should we condition on to understand the relationship between religion and well-being?

We have two sets adjustments sets to chose from. We will pick the set with the measured indicators.

@@

### Note: there is collider counfounding if we include all variables into our model

@@

## Specifying the mechanism in relation to this dataset.

@@

There are many theories of how religion affects well-being we might consider (cite Highland et al). We will focus on two non-trivial theories -- one related to religious behavior and another related to religious belief.  (Citations to follow ...)

@@

### Dag shoes two adjustment sets
```{r}
tidy_ggdag2 <- dagify(
  wb  ~ support + ses + male + country + U2,
  support ~ church + male + country,
  church ~ age + male + country + U1,
  ses ~ country + age + U2,
  exposure =  "church",
  outcome =   "wb"
) %>%
  tidy_dagitty()

@@

### Confounder bias if we include our theoretically proposed causal mediators!

Before proceeding with the model, note a curiosity. If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.

@@

Let this serve as a warning for us to think about our causation. We must develop our causal model before we leap into our statistical model.

## The statistical model

@@

## degroup variables for within and between analysis
## https://cran.r-project.org/web/packages/parameters/vignettes/demean.html
d1 <-
  cbind(d1,
        demean(d1, select = c("anx", "ch", "bg"), group = "country"))
```

@@

We can see there is no effect of church attendance on anxiety (according to this model). (Note that throughout, the plots will sample 300 draws from the posterior distribution. )

@@

Before racing ahead, lets do some posterior predictive checks on the model

The trace plots look good.

@@

However the posterior predictive check reveals that we should work with the original untransformed values of the outcome variable data.

@@

or we can simulate and see that our prior is scattered across five standard deviations of anxiety, and we can see our prior is telling us that anything can happen
.

@@

## The group level effects of country

There is lots of variability at the country level intercepts for anxiety. We can see that India is over .5 standard deviations  less anxious than the average culture level anxiety. By contrast, Japan, Israel, and France are over .3 standard deviations more anxious than average. There looks to be perhaps some covariance between the country intercepts and slopes... maybe though all estimates crosses zero. We'd either want to use stronger priors or collect data from more countries than are in this sample.

@@

Let's delving into this question of country level effects by asking whether people who are *relatively* more church attending within their countries tending to have more anxiety.

@@

- between_church centering = country level differences in church attendance.

@@

The model.

@@

We find no relation of "ch_within" to anxiety.

@@

However, a strong between-country effect emerges:

@@

Before considering this, let's run some model checks workflow as before.
